Harnessing the cloud to understand light: A whitepaper detailing the \MaxwellFDS core solver

### 2013-04-16 thesis/appendix2.tex
Draft done.

\chapter[Design of optical cloaks and mimics]{Objective-first design of optical cloaks and mimics}\label{ob-1 meta}
------------------------------------------------------%
\section{Optical Cloak Design}
%---------------------------------------------------------%    
We now extend the applicability of our method
    to the design of metamaterial devices which operate in free-space.
In particular,
    we adapt the waveguide coupler algorithm to the 
    to the design of optical cloaks.

>>>>

\chapter[Design of optical cloaks and mimics]{Objective-first design of optical cloaks and mimics}\label{ob-1 meta}
% ------------------------------------------------------%
\section{Optical Cloak Design}
%---------------------------------------------------------%    
We now further extend the applicability of the objective-first method
    (chapter~\ref{ob-1})
    to the design of metamaterial devices which operate in free-space.
In particular,
    we adapt the waveguide coupler algorithm to the 
    to the design of optical cloaks.

### 2013-04-16 thesis/chapter1.tex
Draft done.

\section{Problem formulation}
This work was originally sparked by the realization that 
    the electromagnetic wave equation was not only linear in the field variable,
    but also in the structure variable.
Specifically, that
    \BE \nabla \times \mu_0^{-1} \nabla \times E - \omega^2 \epsilon E = 
        -i\omega J \label{p1:wave} \EE
    is \emph{bi-linear} in both $E$ and $\epsilon$.

We now make the following substitution in order to represent \eq{p1:wave}
    in linear algebra notation:
    \BI $E \to x$, 
    \I  $\epsilon \to z$,
    \I  $\nabla\times\mu_0^{-1}\nabla\times - \omega^2 \epsilon \to A(z)$, and
    \I  $-i\omega J \to b$. \EI
The electromagnetic wave equation (\eq{p1:wave}) can now be expressed simply as
    \BE A(z)x = b, \label{p1:axb} \EE
    which is not only linear in $x$, but also in $z$, 
    since $A(z)$ is linear with respect to $z$.

@@

In the language of physics, this amounts to computing
    \BE \epsilon = (\nabla\times\mu_0^{-1}\nabla\times E + 
        i \omega J)/ \omega^2 E \EE
    given $E$.

In the language of linear algebra, we have formulated the following problem:
    \BE \minimize_z \|A(z)x - b\|^2. \label{p1:problem}\EE

@@

\chapter{Regularized direct design}\label{regularized}
\section{Problem formulation}
The simplest way to produce a well-behaved dielectric structure is 
    to add a regularization term to our direct formulation (\eq{p1:axb}),
    \BE \minimize_z \|A(z)x - b\|^2 + \eta \|z - z_0\|^2. \label{reg:problem}\EE
Here, the regularization term $\eta \|z - z_0\|^2$
    serves to keep $z$ from straying too far from $z_0$, which we choose.
The coefficient $\eta$, then, serves as a knob to control the
    strength of the regularization term.

\section{Implementation}
\Eq{reg:problem} can still be computed in one step
    since the regularization term does not add any computational complexity.
Thus the implementation is only a slight modification
    from that of chapter~\ref{direct}.

@@

Critically, fig.~\ref{regls pic} shows that there seems
    to be an inherent trade-off between 
        constraining the structure ($z$) 
        and accurately reproducing the field ($x$).

>>>>

\section{Problem formulation}
This work was originally sparked by the realization that 
    the electromagnetic wave equation was not only linear in the field variable,
    but also in the structure variable.
Specifically, that
    \BE \nabla \times \mu_0^{-1} \nabla \times E - \omega^2 \epsilon E = 
        -i\omega J \label{p1:wave} \EE
    is \emph{bi-linear} in both $E$ and $\epsilon$.
Practically, this meant that solving for $\epsilon$ should be rather straightforward,
    since we already now how to solve linear systems in general.

The language of linear algebra will make itself very useful
    in eventually formulating our problem correctly and generally.
To this end, we now make the following substitution in order to represent \eq{p1:wave}
    in such linear algebra notation:
    \BI $E \to x$, 
    \I  $\epsilon \to z$,
    \I  $\nabla\times\mu_0^{-1}\nabla\times - \omega^2 \epsilon \to A(z)$, and
    \I  $-i\omega J \to b$. \EI
The electromagnetic wave equation (\eq{p1:wave}) can now be expressed simply as
    \BE A(z)x = b. \label{p1:axb} \EE
We note once again that \eq{p1:wave} is not only linear in $x$, but also in $z$, 
    since $A(z)$ is linear with respect to $z$.

@@

In the language of physics, this amounts to a very simple computation;
    namely,
    \BE \epsilon = (\nabla\times\mu_0^{-1}\nabla\times E + 
        i \omega J)/ \omega^2 E \EE
    given $E$!

In the language of linear algebra, the problem is formulation as follows:
    \BE \minimize_z \|A(z)x - b\|^2. \label{p1:problem}\EE

@@

\chapter{Regularized direct design}\label{regularized}
\section{Problem formulation}
We were previously quite excited to learn that solving for 
    a dielectric structure ($\epsilon$ or $z$) given 
    its electromagnetic field ($E$ or $x$) is actually quite simple,
    but generally results in unrealizable structures.
Along these lines, we also knew that the simplest way 
    to produce a well-behaved dielectric structure is 
    to add a regularization term to our direct formulation (\eq{p1:axb}),
    \BE \minimize_z \|A(z)x - b\|^2 + \eta \|z - z_0\|^2. \label{reg:problem}\EE
Here, the regularization term $\eta \|z - z_0\|^2$
    serves to keep $z$ from straying too far from $z_0$, which we choose.
The coefficient $\eta$, then, serves as a knob to control the
    strength of the regularization term.

\section{Implementation}
Fortunately, \eq{reg:problem} can still be computed in one step
    since the regularization term does not add any computational complexity.
Thus the implementation is only a slight modification
    from that of chapter~\ref{direct}.

@@

Although our results seemed promising in producing well-behaved
    dielectric structures;
    critically, fig.~\ref{regls pic} shows that there seems
    to be an unavoidable trade-off between 
    constraining the structure ($z$) 
    and accurately reproducing the field ($x$).

### 2013-04-16 thesis/chapter2.tex
Draft done.

\chapter{Iterative design} \label{iterative}
% \section{Motivation for a Complementary Optimization Strategy}
The fundamental problem in the examples 
    in chapters \ref{direct} and \ref{regularized}
    is actually not in the methods themselves, 
    but in the improper selection of a target field. 
In fact, it is very difficult to select a suitable target field 
    because not every field even has a manufacturable dielectric structure 
    that is able to reproduce it. 
Furthermore, it is nearly impossible to select a multi-dimensional field 
    which corresponds to a well-behaved, isotropic and 
    discretely-valued $\epsilon$, 
    as would be needed for practical structures.

@@

\section{Problem formulation}
We start with the same target field as in the previous examples 
    but we now allow for it to be modified during the design process. 
Specifically, we alternate between modifying $z$ (structure) 
    to better fit $x$ (field), 
    and then modifying the $x$ (field) to better fit $z$ (structure):
    \begin{subequations}\begin{align} 
    \minimize_z & \|A(z)x - b\|^2 + \eta_0\|z - z_\text{prev}\|^2 \\
    \minimize_x & \|A(z)x - b\|^2 + \eta_1\|x - x_\text{prev}\|^2,
    \end{align}\end{subequations}
which can be more succintly expressed as 
    \BE \minimize_\text{alternately $x$ then $z$} 
        \|A(z)x - b\|^2 + \eta_0\|z - z_\text{prev}\|^2
                        + \eta_1\|x-x_\text{prev}\|^2. \label{iterative problem}\EE

@@

\section{Implementation}
The most notable implementation difference is that,
    in contrast to direct design,
    \eq{iterative problem} must be solved multiple times
    in order to arrive at a good design.
Thus, design times now increase from seconds to minutes.

Note that, in order to design two-dimensional structures,
    \eq{iterative problem} is only modified in the
    definition of its constituent vectors and matrices,
    the formulation remains the same.

\section{Results}
\subsection{1D}
Fig.~\ref{comp pic} shows that the iterative algorithm, after $400$ iterations and with the correct choice of regularization parameters $\eta_0$ and $\eta_1$, results in a well-behaved structure that is able to closely reproduce the modified target field\cite{Lu10}.

@@

\chapter{Iterative design with bounded $\epsilon$} \label{bounded}
\section{Problem formulation}
In order to achieve a more practical, discretely-valued dielectric structure, we can impose strict upper- and lower-bounds on the structure ($z$). 
To this end, we can modify \eq{iterative problem} as such,
    \begin{subequations}
    \begin{align} 
    \minimize & \|A(z)x - b\|^2 + \eta_1\|x - x_\text{prev}\|^2\\
    \subto & z_\text{min} \le z \le z_\text{max}. 
    \end{align}\label{bounded problem}
    \end{subequations}
Here, the regularization term for $z$ has been replaced by hard constraints
    on the values that $z$ can take on.

This modification takes us one step closer to producing manufacturable structures,
    which must be discretely valued.

\section{Implementation} 
In this algorithm, \eq{bounded problem} is no longer linear in $z$
    but is still convex\cite{Boyd04}.
For this reason, the \emph{CVX} package\cite{Grant09}, 
    a Matlab-based modeling system for convex optimization, 
    is now used to solve \eq{bounded problem} for $z$.

>>>>

\chapter{Iterative design} \label{iterative}
% \section{Motivation for a Complementary Optimization Strategy}
Having started off with the exciting realization that one could easily 
    solve for $z$ ($\epsilon$), we quickly ran into 
    the problem revealed in fig.~\ref{regls pic}.
It now seemed that a structure that was both well-behaved and high-performance
    might not be possible!

Further thought led us to conclude that
    the fundamental problem in the examples 
    in chapters \ref{direct} and \ref{regularized}
    is actually not in the methods themselves, 
    but in the improper selection of a target field. 
In fact, we realizaed thtat it is very difficult to select a suitable target field 
    because not every field even has a manufacturable dielectric structure 
    that is able to reproduce it. 
Furthermore, it is nearly impossible to select a multi-dimensional field 
    which corresponds to a well-behaved, isotropic and 
    discretely-valued $\epsilon$, 
    as would be needed for practical structures.

@@

\section{Problem formulation}
The problem formulation for our iterative design strategy
    starts with the same target field as in the previous examples,
    but we now allow for it to be modified during the design process. 
Specifically, we alternate between modifying $z$ (structure) 
    to better fit $x$ (field), 
    and then modifying the $x$ (field) to better fit $z$ (structure):
    \begin{subequations}\begin{align} 
    \minimize_z & \|A(z)x - b\|^2 + \eta_0\|z - z_\text{prev}\|^2 \\
    \minimize_x & \|A(z)x - b\|^2 + \eta_1\|x - x_\text{prev}\|^2,
    \end{align}\end{subequations}
which can be more succintly expressed as 
    \BE \minimize_\text{alternately $x$ then $z$} 
        \|A(z)x - b\|^2 + \eta_0\|z - z_\text{prev}\|^2
                        + \eta_1\|x-x_\text{prev}\|^2. \label{iterative problem}\EE

@@

\section{Implementation}
In terms of implementation,
    the most notable implementation difference is that,
    in contrast to direct design,
    \eq{iterative problem} must be solved multiple times
    in order to arrive at a good design.
Thus, design times now increase from seconds to minutes.

Since this strategy started to produce more successful results,
    we shifted from one-dimensional structures to two-dimensional structures.
In terms of implementation, we note that 
    in order to design higher-dimensional structures,
    \eq{iterative problem} is only modified in the
    definition of its constituent vectors and matrices,
    the formulation remains the same.

\section{Results}
\subsection{1D}
Fig.~\ref{comp pic} shows that the iterative algorithm, after $400$ iterations and with the correct choice of regularization parameters $\eta_0$ and $\eta_1$, 
    indeed results in a well-behaved structure that is able to closely reproduce the modified target field\cite{Lu10}!

@@

\chapter{Iterative design with bounded $\epsilon$} \label{bounded}
\section{Problem formulation}
Having achieved a measure of success with the iterative design strategy
    outlined in the previous chapter,
    we strove 
    to achieve a more practical, discretely-valued dielectric structure. 
To do this, we imposed strict upper- and lower-bounds on the structure variable $z$. 
To this end, we modified \eq{iterative problem} as such,
    \begin{subequations}
    \begin{align} 
    \minimize & \|A(z)x - b\|^2 + \eta_1\|x - x_\text{prev}\|^2\\
    \subto & z_\text{min} \le z \le z_\text{max}. 
    \end{align}\label{bounded problem}
    \end{subequations}
Here, the regularization term for $z$ has been replaced by hard constraints
    on the values that $z$ can take on.

\section{Implementation} 
In this formulation, \eq{bounded problem} is no longer linear in $z$
    but is still convex\cite{Boyd04}.

Practically, this means that although it is no longer absolutely trivial
    to solve \eq{bounded problem} for $z$, it can still be done reliably.
By reliably, we mean that since \eq{bounded problem} is still convex in $z$,
    there exists a single local optimum (which is also the global optimum)
    and that there also exists software which can efficiently
    discover this globally optimum point.

Specifically, the \emph{CVX} package\cite{Grant09}, 
    a Matlab-based modeling system for convex optimization, 
    is the software we used to solve \eq{bounded problem} for $z$.

### 2013-04-16 thesis/intro.tex
Draft done.

\chapter{Introduction}
And now for an introduction...

>>>>

\chapter{Introduction}
The twenty-first century has rightly been called the ``information age'',
    as advances in computation have resulted in an explosion
    in the amount of information created and communicated on a daily basis.

The underlying communication fabric and key technological enabler
    of our information age has been the high-bandwidth, long-distance
    communication capability afforded by the optical network.

First used to communicate between cities and across continents,
    optical networks are now needed to simply communicate across rooms (datacenters)
    because of the ballooning information creation and communication
    abilities of single computers.
As our appetite for information continues to grow,
    the optical network will be used for even shorter distances
    such as chip-to-chip and even core-to-core communication.
The focus of this work is the design of nanophotonic components
    to enable such future on-chip optical networks.

Currently, on-chip networks consist of devices
    which are designed by tuning
    a handful of design parameters (\fig{integrated_circuit}).
Much progress has been made on this front;
    however, we note that the available design space is largely untouched (\fig{des_complexity}).

\myfig{integrated_circuit}
    {Examples of hand-tuned nanophotonic devices.
    Device performance is tuned by varying the sizes of and spacing between
    a handful of elements (waveguides, resonators, gratings, \ldots).}

To be precise, the motivation for this works stems from the twin realizations that
    \begin{enumerate}
    \item the fabrication of on-chip optical networks in semiconductor 
    foundries enables virtually unlimited design complexity at no additional 
    fabrication cost, and
    \item component performance can only benefit by the addition and
    consideration of these additional degrees of freedom.
    \end{enumerate}
The goal of this work, then, is to be able to take advantage
    of the full available parameter space to design nanophotonic devices.

\myfig{des_complexity}
    {The parameter space for nanophotonic design is enormous and largely untapped.
    Here, we show that, 
        limited to including or excluding 100 nm square areas,
        a 1500 nm $\times$ 1500 nm area already contains $2^{225}$ 
        (an uncountable number) possible designs.}

Along this line of reasoning
    we expect to be able to produce devices which
    \BI have very compact footprints,
    \I  exhibit high efficiency performance,
    \I  are multi-modal in their operation,
    \I  are robust to shifts in wavelength and temperature,
    \I  are robust to fabrication error,
    \I  may even demonstrate novel, never-before-seen functionality. \EI
Our final result (chapter~\ref{final}) indeed demonstrates
    manufacturable, three-dimensional nanophotonic devices which
    exhibit these properties.

Of course, in this regime, the hand-tuning of parameters 
    is no longer a tractable strategy.
Instead, we use a \emph{design-by-specification} strategy
    which simply means that our software produces nanophotonic designs
    based solely based on the desired performance specification given to it,
    and requires no intermediate hand tuning.

We note, lastly, that the design-by-specification strategy
    is unusually powerful in the sense that \emph{all} useful nanophotonic devices
    can be designed in this way.
Indeed, the apparent ease with which we are able to produce designs seems
    to lend one to this conclusion.

The organization of this thesis is straightforward: 
    chapters \ref{direct}-\ref{ob-1} detail the incremental steps in achieving
    our final result, which is presented in chapter \ref{final}.
Futhermore, each section is broken into three major sections:
    \begin{description}
    \item[Problem formulation] where we show how we cast the design problem
        in incrementally more sophisticated mathematics.
    \item[Implementation] where we note how the formulated problem 
        was solved computationally.
    \item[Results] where we demonstrate the abilities (and weaknesses)
        of each formulation.
    \end{description}

Additionally, appendices~\ref{ob-1 additional} and \ref{ob-1 meta} 
    contain numerous examples of devices designed using our
    so-called ``objective-first'' design formulation.
Appendix~\ref{maths} contains the nitty-gritty mathematics
    that was used to implement our final design tool, 
    as outlined in chapter~\ref{final}.
This appendix will be most useful to any wishing to reproduce our work.
